---
title: "Models Round 2"
author: "Will Jones"
date: "March 11, 2016"
output: html_document
---

```{r load-packages, echo=F, results='hide', message=F}
packages <- c("dplyr", "ggplot2", "ggthemes", "caret", "lubridate", "gamm4",
              "gridExtra", "binom", "tidyr", "knitr", "stargazer", "extrafont",
              "grid", "AUC")
sapply(packages, library, character.only = TRUE)
source("prob-var-plot.R")
source("separation-plot.R")
```


```{r load-data, echo=F}
# Load data
#load('./rides.RData')
#rides_final <- rides %>%
#  mutate(length_meters = length)

rides <- read.csv('sample.csv',
                  col.names = c("trip_id", "rider", "length_meters", 
                                "datetime", "rating", "rating_text"),
                  colClasses = c("character", "factor", "numeric", 
                                 "character", "factor", "factor"))
rides_final <- rides %>%
  mutate(datetime = ymd_hms(datetime, tz="America/Los_Angeles"),
         stressful = ifelse(rating_text == "none", NA,
                            ifelse(rating_text == "good", FALSE, TRUE)),
         rider = as.factor(as.numeric(rider)))

# Small sample for testing
#rides_final <- sample_n(rides_final, 5e3)

###################################
## Comment this out later!!!
# Shift the dates for the test data
#day(rides_final$datetime) <- day(rides_final$datetime) - 200
#rides_final$rider <- as.factor(rep(1:3, length.out = nrow(rides_final)))

load('./weather_daily.RData')
load('./rain_hourly.RData')
```

```{r transform-data, echo=F}
# Join in various data sources
rides_final <- rides_final %>%
  mutate(date = floor_date(datetime, "day")) %>%
  left_join(weather, by = "date") %>%
  mutate(datetime_hour = floor_date(datetime, "hour")) %>%
  left_join(rain, by=c("datetime_hour" = "datetime"))

rides_final <- mutate(rides_final, time = datetime)
day(rides_final$time) <- 1
month(rides_final$time) <- 1
year(rides_final$time) <- 2015
rides_final <- rides_final %>%
  mutate(time = as.numeric(time),
         time = time - min(time),
         time = time / 3600) # convert time from seconds to hours

rides_final$weekend <- as.factor(wday(rides_final$datetime) %in% c(1, 7))

# Filter out the rides with zero length
rides_scaled <- rides_final %>% filter(length_meters != 0)

# Scale the variables
scaled_log_length <- scale(log(rides_scaled$length_meters))
scaled_length <- scale(rides_scaled$length_meters)
scaled_temp <- scale(rides_scaled$mean.temp)
rides_scaled$length <- scaled_length[,1]
rides_scaled$log_length <- scaled_log_length[,1]
rides_scaled$mean.temp <- scaled_temp[,1]
rides_scaled <- rides_scaled %>%
  filter(!is.na(mean.temp) &
         !is.na(rainfall))

rides_scaled <- rides_scaled %>%
  mutate(gust.speed = ifelse(is.na(gust.speed), 0, gust.speed))

# Reset the rider indexes
rides_scaled$rider <- plyr::mapvalues(rides_scaled$rider,
                                      from = levels(rides_scaled$rider),
                                      to = 1:length(levels(rides_scaled$rider)))
```

Log length and temperature were standardized. The center of log length was 
`r attr(scaled_log_length, "scaled:center")` and the scale factor was 
`r attr(scaled_log_length, "scaled:scale")`. The center of temperature was
`r attr(scaled_temp, "scaled:center")` and the scale factor was
`r attr(scaled_temp, "scaled:scale")`.


```{r summary, echo=FALSE}
total_rides <- nrow(rides_scaled)
#rides_scaled <- sample_n(rides_scaled, 3000)
# Find riders with at least 20 rated rides
rider_counts <- rides_scaled %>% filter(!is.na(stressful)) %>% group_by(rider) %>% summarise(count = n()) %>%
  filter(count > 20)
rides_scaled <- rides_scaled %>% filter(rider %in% rider_counts$rider)

rated_rides <- sum(!is.na(rides_scaled))
unrated_rides <- 

rides_scaled <- rides_scaled %>% filter(!is.na(stressful))
  
neg_ratings <- sum(rides_scaled$stressful == TRUE)
pos_ratings <- sum(rides_scaled$stressful == FALSE)

num_riders <- rides_scaled$rider %>% unique() %>% length()

summary(rides_scaled$datetime)
```

In this dataset, there are `r total_rides` rides, `r rated_rides / total_rides * 100`
percent of which are rated (`r rated_rides` that were rated and `r unrated_rides`
that were unrated.)

`r neg_ratings / rated_rides * 100` percent of rated rides were rated negatively
(`r neg_ratings` negative ratings and `r pos_ratings` positive ratings.)

There are a total of `r num_riders` riders with more than 20 rides.

```{r rider-plot, echo=F}
rider_avg <- rides_final %>%
  filter(!is.na(stressful)) %>%
  group_by(rider) %>%
  summarise(avg_rating = mean(stressful)) %>%
  arrange(desc(avg_rating)) %>%
  mutate(rider = as.character(rider))

rider_avg_plot <- qplot(x = avg_rating, data = rider_avg, bins = 30) + 
  theme_bw(base_family="CMU Serif") + 
  scale_x_continuous(limits = c(0, 1)) + 
  labs(title = "Distribution of Average Rider Rating")

rider_avg_plot

ggsave(rider_avg_plot, file="plots/rider_avgs.pdf", height = 3, width = 5)
embed_fonts("plots/rider_avgs.pdf", outfile="plots/rider_avgs2.pdf")
```

## Exploratory Plots

```{r make-plots, echo=FALSE, eval=FALSE}
#length_plot <- prob_var_plot(rides_final, "length_meters", "stressful") +
#  theme_bw(base_family="CMU Serif") + 
#  labs(title = "Distribution of Negative Rating by Length",
#       x = "length of ride", y = "prob. of negative rating",
#       window_width = 20, rug = FALSE)

#length_plot
#
#m <- gam(stressful ~ s(length_meters), data = rides_final, family = binomial)
#plot(m)

#ggsave(length_plot, file="plots/length_prob_plot.pdf", width = 5, height = 3)

hour_plot <- prob_var_plot(rides_final, "time", "stressful") +
  theme_bw(base_family="CMU Serif") + 
  labs(title = "Distribution of Negative Rating by Time of Day",
       x = "hour of day", y = "prob. of negative rating") +
  scale_x_continuous(breaks = seq(0, 24, 3))

hour_plot

m <- gam(stressful ~ s(time, bs="cc", by=weekend), data = rides_final, family = binomial)
plot(m)

ggsave(hour_plot, file = "plots/hour_prob_plot.pdf", height = 3, width = 5)
```

```{r time-length-plot}
rides_scaled %>%
  filter(weekend == "FALSE") %>%
ggplot(aes(x = time, y = log_length)) + 
    geom_point(alpha = 0.4) +
  #geom_density_2d() +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = 0.3, color = NA) +
    scale_x_continuous(breaks = c(0, 6, 12, 18, 24),
                     labels = c("12am", "6am", "12pm", "6pm", "12am")) + 
  theme_bw()
```


## Models

```{r fit-models}
# Baseline: Logistic Regression
model1 <- glm(stressful ~ log_length + mean.temp + gust.speed + rainfall + rainfall.4h + 
                mean.wind.speed,
              data = rides_scaled, family=binomial)
# Add rider intercepts
model2 <- glmer(stressful ~ 1 + (1|rider) + log_length + mean.temp + gust.speed + 
                  rainfall + rainfall.4h+ 
                mean.wind.speed,
              data = rides_scaled, family=binomial)

one_day <- 24 #hours

model3 <- glmer(stressful ~ 1 + (1|rider) + log_length + mean.temp + gust.speed + 
                  rainfall + rainfall.4h + mean.wind.speed +
                  I(sin(2 * pi / one_day * time)):weekend + 
                  I(cos(2 * pi / one_day * time)):weekend + 
                  I(sin(4 * pi / one_day * time)):weekend + 
                  I(cos(4 * pi / one_day * time)):weekend,
              data = rides_scaled, family=binomial)

# Add smoother for time of day
model4 <- gamm4(stressful ~ 1 + log_length + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h  + s(time, bs="cc", k=9, by = weekend),
                random =~(1|rider),
                knots=list(time=seq(0, 24, 3)),
                data = rides_scaled, family = binomial)

# Add smoother for length
model5 <- gamm4(stressful ~ 1 +  s(log_length) + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h +
                  s(time, bs="cc", k=9, by = weekend),
                random =~(1|rider),
                knots=list(time=seq(0, 24, 3)),
                data = rides_scaled, family = binomial)

# Remove rider intercepts
model6 <- gam(stressful ~ 1 + log_length + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h + 
                  s(time, bs="cc", k=9, by = weekend),
                knots=list(time=seq(0, 24, 3)),
                data = rides_scaled, family = binomial)
```


```{r, echo=FALSE}
invlogit <- function (x) { 1 / (1 + exp(-x))}

data_subset <- rides_scaled %>% slice(1:20)
# Trying to get predictions that combine smoothers and random intercept. It's hard.
predict_gamm <- function(model, newdata) {
  intercepts <- coef(model$mer)$rider[["(Intercept)"]]
  intercept_vector <- intercepts[match(newdata$rider, rownames(coef(model$mer)$rider))]
  invlogit(predict(model$gam, newdata = newdata) + intercept_vector)
}
predict_gamm(model5, data_subset)


```

## Model Fits

```{r training-sep-plots, warning=FALSE}
fitted_values <- data.frame("actual" = rides_scaled$stressful,
                            "predict1" = predict(model1, type="response"),
                            "predict2" = predict(model2, type="response"),
                            "predict3" = predict(model3, type="response"),
                            "predict4" = predict_gamm(model4, rides_scaled),
                            "predict5" = predict_gamm(model5, rides_scaled),
                            "predict6" = predict(model6, type="response"))


for (i in 1:6) {
  p <- separation_plot(fitted_values, "actual", paste("predict", as.character(i), sep = ""))
  ggsave(p, file=paste("plots/model", as.character(i), "-sep.pdf", sep=""), 
         width = 4, height = 0.25)
}

#ggsave(train_sep_plots, file="plots/sep_plots.pdf",
#       width = 5, height = 3)

prob_eval_plot <- function(df, actual_col, pred_col, title) {
  prob_var_plot(df, pred_col, actual_col, window_width = 0.03, rug = FALSE) + 
  theme_bw(base_family="CMU Serif") + 
    geom_abline(slope = 1, intercept = 0, color = "red") + 
    scale_x_continuous(limits=c(0,1)) + 
    labs(title = title, y = "Empirical Probability",
         x = "Predicted Probability")
}

prob_eval_plots <- grid.arrange(
  prob_eval_plot(fitted_values, "actual", "predict1", "Model 1"),
  prob_eval_plot(fitted_values, "actual", "predict2", "Model 2"),
  prob_eval_plot(fitted_values, "actual", "predict3", "Model 3"),
  prob_eval_plot(fitted_values, "actual", "predict4", "Model 4"),
    prob_eval_plot(fitted_values, "actual", "predict5", "Model 5"),
    prob_eval_plot(fitted_values, "actual", "predict6", "Model 6"),
  ncol = 2
)

ggsave(prob_eval_plots, file="plots/prob_eval_plots.pdf",
       width = 6, height = 6)
```


```{r performance-measures}
# Loglikelihood
logLik(model1)
logLik(model2)
logLik(model3)
logLik(model4$mer)
logLik(model5$mer)
logLik(model6)

# AIC
model1$aic
AIC(model2)
AIC(model3)
AIC(model4$mer)
AIC(model5$mer)
AIC(model6)

# AUC
calc_auc <- function(model) auc(sensitivity(fitted_values[[model]], as.factor(as.numeric(fitted_values$actual))))
calc_auc("predict1")
calc_auc("predict2")
calc_auc("predict3")
calc_auc("predict4")
calc_auc("predict5")
calc_auc("predict6")
```

```{r summary-table, results='asis', echo=F}
stargazer(model1, model2, model3, model6, type="html", title="results",
          align=TRUE, ci = TRUE)
```


```{r model-confint}
kable(confint.lm(model1))
kable(confint(model2, method="Wald"))
kable(confint(model3, method="Wald"))
kable(confint(model4$gam))
kable(confint(model5$gam))
```

```{r gam-model-summaries}
summary(model4$gam)
summary(model5$gam)
```


## Model Comparison

```{r}
anova(model2, model3)
```


```{r cv-split}

folds <- createFolds(rides_scaled$rider)

compute_auc <- function(predicted, actual) auc(sensitivity(predicted, as.factor(as.numeric(actual))))

one_auc <- function(fold, fit)  {
  the_model <- rides_scaled[-fold,] %>% fit() 
  if (typeof(the_model) == "S4" || is.null(the_model$gam)){
    predictions <- the_model %>%  predict(newdata = rides_scaled[fold,])
  }
  else {
    predictions <- predict_gamm(the_model, rides_scaled[fold,])
  }
  compute_auc(predictions, rides_scaled[fold,]$stressful)
}

cv_auc <- function(fit) lapply(folds, one_auc, fit = fit) %>% unlist() %>% mean()

```


```{r refit-models}
# Baseline: Logistic Regression
# model1_t <- glm(stressful ~ log_length + mean.temp + gust.speed + rainfall + rainfall.4h + weekend,
#               data = rides_train, family=binomial)
fit1 <- function(data) {
  glm(stressful ~ log_length + mean.temp + gust.speed + rainfall + rainfall.4h + weekend + mean.wind.speed,
              data = data, family=binomial)
}
# Add rider intercepts
# model2_t <- glmer(stressful ~ 1 + (1|rider) + log_length + weekend + mean.temp + gust.speed + 
#                   rainfall + rainfall.4h,
#               data = rides_train, family=binomial)
fit2 <- function(data) {
  glmer(stressful ~ 1 + (1|rider) + log_length + weekend + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h,
              data = data, family=binomial)
}

# Add mean ride length as predictor
#model3_t <- glmer(stressful ~ 1 + (1|rider) + log_length + mean.temp + gust.speed + 
#                  rainfall + rainfall.4h + avg_log_length + weekend,
#              data = rides_train, family=binomial)

# Add sinusoidal time of day
one_day <- 24 #hours

# model3_t <- glmer(stressful ~ 1 + (1|rider) + log_length + mean.temp + gust.speed + 
#                   rainfall + rainfall.4h  + 
#                     I(sin(2 * pi / one_day * time)):weekend + 
#                   I(cos(2 * pi / one_day * time)):weekend + 
#                   I(sin(4 * pi / one_day * time)):weekend + 
#                   I(cos(4 * pi / one_day * time)):weekend,
#               data = rides_train, family=binomial)

fit3 <- function(data) {
  glmer(stressful ~ 1 + (1|rider) + log_length + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h  + 
                    I(sin(2 * pi / one_day * time)):weekend + 
                  I(cos(2 * pi / one_day * time)):weekend + 
                  I(sin(4 * pi / one_day * time)):weekend + 
                  I(cos(4 * pi / one_day * time)):weekend,
              data = data, family=binomial)
}

# Add smoother for time of day
# model4_t <- gamm4(stressful ~ 1 + log_length + mean.temp + gust.speed +
#                   rainfall + rainfall.4h + s(time, bs="cc", k=9, by = weekend),
#                 random =~(1|rider),
#                 knots=list(time=seq(0, 24, 3)),
#                 data = rides_train, family = binomial)

fit4 <- function(data) {
  gamm4(stressful ~ 1 + log_length + mean.temp + gust.speed +mean.wind.speed +
                  rainfall + rainfall.4h + s(time, bs="cc", k=9, by = weekend),
                random =~(1|rider),
                knots=list(time=seq(0, 24, 3)),
                data = data, family = binomial)
}

# Add smoother for length
# model5_t <- gamm4(stressful ~ 1 +  s(log_length) + mean.temp + gust.speed +
#                   rainfall + rainfall.4h +
#                   s(time, bs="cc", k=9, by = weekend),
#                 random =~(1|rider),
#                 knots=list(time=seq(0, 24, 3)),
#                 data = rides_train, family = binomial)

fit5 <- function(data) {
  gamm4(stressful ~ 1 +  s(log_length) + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h +
                  s(time, bs="cc", k=9, by = weekend),
                random =~(1|rider),
                knots=list(time=seq(0, 24, 3)),
                data = data, family = binomial)
}

# model6_t <- gam(stressful ~ 1 + log_length + mean.temp + gust.speed +
#                   rainfall + rainfall.4h +
#                   s(time, bs="cc", k=9, by = weekend),
#                 knots=list(time=seq(0, 24, 3)),
#                 data = rides_train, family = binomial)

fit6 <- function(data) {
  gam(stressful ~ 1 + log_length + mean.temp + gust.speed + mean.wind.speed +
                  rainfall + rainfall.4h +
                  s(time, bs="cc", k=9, by = weekend),
                knots=list(time=seq(0, 24, 3)),
                data = data, family = binomial)
}

cv_auc(fit1)
cv_auc(fit2)
cv_auc(fit3)
cv_auc(fit4)
cv_auc(fit5)
cv_auc(fit6)
```


```{r test-plots, eval = FALSE}
fitted_values_t <- data.frame("actual" = rides_test$stressful,
                            "predict1" = predict(model1_t, rides_test, type="response"),
                            "predict2" = predict(model2_t, rides_test, type="response"),
                            "predict3" = predict(model3_t, rides_test, type="response"),
                            "predict4" = predict_gamm(model4_t, rides_test),
                            "predict5" = predict_gamm(model5_t, rides_test),
                            "predict6" = predict(model6_t, rides_test, type="response"))

test_sep_plots <- grid.arrange(separation_plot(fitted_values_t, "actual", "predict1"),
            separation_plot(fitted_values_t, "actual", "predict2"),
            separation_plot(fitted_values_t, "actual", "predict3"),
            separation_plot(fitted_values_t, "actual", "predict4"),
            separation_plot(fitted_values_t, "actual", "predict5"),
            separation_plot(fitted_values_t, "actual", "predict6"),
            ncol=1)

ggsave(test_sep_plots, file = "plots/sep_plots_out_of_sample.pdf",
       width = 6, height = 6)

test_eval_plots <- grid.arrange(
  prob_eval_plot(fitted_values_t, "actual", "predict1", "Model 1"),
  prob_eval_plot(fitted_values_t, "actual", "predict2", "Model 2"),
  prob_eval_plot(fitted_values_t, "actual", "predict3", "Model 3"),
  prob_eval_plot(fitted_values_t, "actual", "predict4", "Model 4"),
    prob_eval_plot(fitted_values_t, "actual", "predict5", "Model 5"),
    prob_eval_plot(fitted_values_t, "actual", "predict6", "Model 6"),
  ncol = 2
)

ggsave(test_eval_plots, file="plots/eval_plot_out_of_sample.pdf",
       width = 6, height = 6)
```


### Rider Intercepts

```{r}
get_intercepts <- function(model, group_predictors) {
  if (missing(group_predictors)) return(coef(model)$rider[["(Intercept)"]])
  coef_table <- coef(model)$rider
  coef_table$rider <- rownames(coef_table)
  coef_table <- coef_table[,c("rider", "(Intercept)", group_predictors)]
  rider_x <- rides_scaled[,c("rider", group_predictors)] %>% data.frame() %>% distinct()
  coef_table <- left_join(coef_table, rider_x, by="rider")
  intercept <- coef_table[["(Intercept)"]]
  for (i in 1:length(group_predictors)) {
    intercept <- intercept + coef_table[[paste(group_predictors[i], "x", sep=".")]] * coef_table[[paste(group_predictors[i], "y", sep=".")]]
  }
  intercept
}

rider_intercepts <- data.frame("rider" = rownames(coef(model3)$rider[1]),
                             "model2" = get_intercepts(model2),
                             "model4" = get_intercepts(model3),
                              "model5" = coef(model4$mer)$rider[["(Intercept)"]],
                              "model6" = coef(model5$mer)$rider[["(Intercept)"]]) %>%
  tbl_df() %>%
  gather(model, intercept, 2:5)

rider_intercept_avgs <- rider_intercepts %>%
  group_by(model) %>%
  summarise(avg_intercept = mean(intercept))

rider_intercept_plot <- ggplot(rider_intercepts, aes(x = intercept)) + geom_histogram(bins=30) +
  facet_wrap(~ model, ncol = 2) + 
  theme_bw(base_family="CMU Serif") + 
  geom_vline(data = rider_intercept_avgs, size = 1,
             aes(xintercept = avg_intercept), color = "steelblue") + 
  labs(title="Distribution of Rider Intercepts")

rider_intercept_plot

ggsave(rider_intercept_plot, file="plots/rider_intercept_plot.pdf",
       width = 6, height = 4)

# subtract averages
rider_intercepts_normalized <- left_join(rider_intercepts, rider_intercept_avgs, by="model") %>%
  mutate(intercept = intercept - avg_intercept) %>%
  filter(model != "model1")

rider_intercepts_change <- ggplot(rider_intercepts_normalized, aes(x = model, y = intercept, group = rider)) +
  geom_line(alpha = 0.5, size = 0.5) + 
  theme_bw(base_family="CMU Serif") + 
  labs(y = expression(alpha[j]),
       title = "Changes in Intercepts by Model")
ggsave(rider_intercepts_change, file="plots/rider_intercepts_change.pdf",
       width = 6, height = 4)
```

### Time of Day

```{r time-of-day-plots}
gamm_interval <- function(model, newdata) {
  intercepts <- coef(model$mer)$rider[["(Intercept)"]]
  predictions <- predict(model$gam, newdata = newdata, se.fit = TRUE) %>% 
    data.frame() %>%
    mutate(lwr = fit - 2 * se.fit, upr = fit + 2 * se.fit) %>%
    mutate(se.fit = NULL) %>%
    tbl_df()
  
  predictions$intercept <- intercepts[match(newdata$rider, rownames(coef(model$mer)$rider))]
  
  predictions <- predictions %>%
    mutate(fit = invlogit(fit + intercept),
           upr = invlogit(upr + intercept),
           lwr = invlogit(lwr + intercept)) %>%
    mutate(intercept = NULL)
  
  predictions
}


# Want to find a rider ID that seems most typical
intercepts <- coef(model5$mer)$rider[["(Intercept)"]]
rider_labels <- levels(as.factor(as.numeric(rides_scaled$rider)))
typicalest_rider <- rider_labels[which.min(intercepts - mean(intercepts))]

res <- 200
  
fake_time_data <- data.frame("time" = rep(seq(0, 24, length.out = res),2),
                             "rider" = as.factor(rep(typicalest_rider, 2*res)),
                             "log_length" = rep(median(rides_scaled$log_length), 2*res),
                             "mean.temp" = rep(mean(rides_scaled$mean.temp), 2*res),
                             "mean.wind.speed" = rep(0, 2*res),
                             "gust.speed" = rep(0, 2*res),
                             "rainfall" = rep(0, 2*res),
                             "rainfall.4h" = rep(0, 2*res),
                             "weekend" = as.factor(c(rep(TRUE, res), rep(FALSE, res))))

fake_time_data <- fake_time_data %>%
  mutate(model3_fit = predict(model3, newdata = fake_time_data, type="response"),
         model3_upr = model3_fit-1000,
         model3_lwr = model3_fit-1200)
fake_time_data <- gamm_interval(model4, fake_time_data) %>%
  dplyr::rename(model4_fit = fit, model4_upr = upr, model4_lwr = lwr) %>%
  cbind(fake_time_data) %>%
  tbl_df()
fake_time_data <- gamm_interval(model5, fake_time_data) %>%
  dplyr::rename(model5_fit = fit, model5_upr = upr, model5_lwr = lwr) %>%
  cbind(fake_time_data) %>%
  tbl_df()
fake_time_data <- predict(model6, fake_time_data, se.fit = TRUE, type = "response") %>%
  data.frame() %>% tbl_df() %>%
  dplyr::mutate(lwr = fit - 2*se.fit, upr = fit + 2*se.fit, se.fit = NULL) %>%
  dplyr::rename(model6_fit = fit, model6_upr = upr, model6_lwr = lwr) %>%
  cbind(fake_time_data) %>%
  tbl_df()

#predictions_from_1 <- predict(model1, type="response")

fake_time_data <- fake_time_data %>%
  gather(model, prediction, 1:9, model3_fit, model3_upr, model3_lwr) %>%
extract(model, c("model", "part"), regex = "([:alnum:]+)_([:alnum:]+)") %>%
  distinct() %>%
  spread(part, prediction)

fake_time_data$weekend <- plyr::mapvalues(fake_time_data$weekend,
                                      from = c(FALSE, TRUE),
                                      to = c("weekday", "weekend"))

time_plot <- filter(fake_time_data, model != "model6") %>%
  ggplot(aes(x = time, y = fit, color = model)) +
  geom_line() + 
    geom_line(aes(y = lwr),
              linetype = "dotted") +
    geom_line(aes(y = upr),
              linetype = "dotted") +
  facet_wrap(~ weekend, nrow = 1) +
  theme_bw(base_family="CMU Serif") + 
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24),
                     labels = c("12am", "6am", "12pm", "6pm", "12am"))  + 
  scale_y_continuous(limits = c(0, 0.007)) + 
  labs(x = "hour of day",
       y = "p(negative rating)",
       title = "Fitted Model Functional Forms for Time of Day") + 
  theme(legend.position="bottom")

#qplot(x =rides_scaled$time, y = predictions_from_2, alpha = 0.25)

time_plot

ggsave(time_plot, file="plots/time_fit_plot.pdf",
       width = 6, height = 3)

time_plot_6 <- filter(fake_time_data, model == "model6") %>%
  ggplot(aes(x = time, y = fit)) +
  geom_line() + 
      geom_line(aes(y = lwr),
              linetype = "dotted") +
    geom_line(aes(y = upr),
              linetype = "dotted") +
  facet_wrap(~ weekend, nrow = 1) +
  theme_bw(base_family="CMU Serif") + 
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24),
                     labels = c("12am", "6am", "12pm", "6pm", "12am"))  +
  scale_color_grey() +
  labs(x = "hour of day",
       y = "p(negative rating)",
       title = "Time of Day Fit for Model 6") 

ggsave(time_plot_6, file="plots/time_fit_plot_6.pdf",
       width = 6, height = 2)
```

Weird. It seems the random intercepts allow us to encode a lot of information
really quickly.

## Random Intercept Tests

```{r}

model_riders <- glmer(stressful ~ 1 + (1|rider) + log_length + weekend + mean.temp + gust.speed + 
                  rainfall + rainfall.4h,
              data = rides_scaled, family=binomial)

rides_scaled$fake_rider <- sample(1:num_riders, nrow(rides_scaled), replace = TRUE)

model_random <- glmer(stressful ~ 1 + (1|fake_rider) + log_length + weekend + mean.temp + gust.speed + 
                  rainfall + rainfall.4h,
              data = rides_scaled, family=binomial)

intercept_test_fitted_values <- data.frame("actual" = rides_scaled$stressful,
                            "predict_rider" = predict(model_riders, type="response"),
                            "predict_random" = predict(model_random, type="response"))

intercept_test_sep_plots <- grid.arrange(
  separation_plot(intercept_test_fitted_values, "actual", "predict_rider"),
  separation_plot(intercept_test_fitted_values, "actual", "predict_random"),
  ncol=1, padding = unit(c(0, 0, 0, 0), "cm"))

intercept_test_sep_plots
ggsave(intercept_test_sep_plots, file="plots/intercept_test_plot.pdf",
       width = 5.5, height = 1)
```


```{r}
time_pred <- fitted_values
time_pred$time <- rides_scaled$time
time_pred$weekend <- rides_scaled$weekend
time_pred <-  time_pred %>%
  gather(model, pred, 2:7) %>%
  mutate(model = factor(model))
time_pred$model <- plyr::mapvalues(time_pred$model,
                                   from = levels(time_pred$model),
                                   to = c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5", "Model 6"))

time_pred_plot <- ggplot(filter(time_pred, weekend == FALSE), aes(x = time, y = pred)) +
  geom_point(alpha = 0.1, size = 0.4) + 
  facet_wrap(~model) +
  theme_bw(base_family="CMU Serif") + 
  labs(title = "Model Predictions by Time")


time_pred_plot

ggsave(time_pred_plot, file="plots/time_pred_plot.pdf",
       width = 6, height = 4)
```

# What do the rider intercepts tell us?

Let's try and figure out the probability that each rider would give a typical
ride a negative rating.

```{r}
res <- 200
num_riders <- rides_scaled$rider %>% unique() %>% length()
fake_rider_data <- data.frame("rider" = unique(rides_scaled$rider),
                              "time" = rep(12, num_riders),
                             "log_length" = rep(median(rides_scaled$log_length), num_riders),
                             "mean.temp" = rep(mean(rides_scaled$mean.temp), num_riders),
                             "gust.speed" = rep(0, num_riders),
                               "mean.wind.speed" = rep(0, num_riders),
                             "rainfall" = rep(0, num_riders),
                             "rainfall.4h" = rep(0, num_riders),
                             "weekend" = as.factor(rep(FALSE, num_riders)))

fake_rider_data$model4 <- predict_gamm(model4, fake_rider_data)

fake_rider_data <- fake_rider_data %>%
  gather(model, prediction, model4)

rider_tendency <- ggplot(fake_rider_data, aes(x = prediction)) + 
  geom_histogram(bins = 20) + 
  theme_bw(base_family="CMU Serif") + 
  labs(x = "p(negative rating)",
       title = "Rider Tendency for Negative Rating") 

ggsave(rider_tendency, file="plots/rider_predictions.pdf",
       width =5, height = 3)
```

