---
title: "First Model"
author: "Will Jones"
date: "December 8, 2015"
output: 
  html_document: 
    keep_md: yes
---

```{r load-packages, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(arm))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(stargazer))
```


## Introduction

As our first steps in modeling ride rating, we will start to model without route
data. Instead we will focus on other question in the modeling as a start for our
model:

- How much variation is there between riders in how they tend to rate rides?
- What relationship does weather, like rain or wind speed, have with ride
rating?
- How does ride rating fluctuate with time of day (which we use as a proxy for
traffic)?

We actually expect a fair amount of the variance in ride rating to be explained
by these variables, based on tests of a smaller sample.

## Some Numbers about the Data

```{r load_data, echo=FALSE}
load('./rides.RData')
load('./weather_daily.RData')
load('./rain_hourly.RData')
```

```{r summarize_data, echo=FALSE}
total.obs <- nrow(rides)
no.response <- rides %>% filter(is.na(stressful)) %>% nrow()
```

There are `r total.obs` rides in the data set, with `r no.response`
(`r 100 * no.response / total.obs`%) rides with no rating.

## What variables will we include?

```{r merge_data, echo=FALSE}
# Merging in weather data
rides.final <- rides %>%
  mutate(date = floor_date(datetime, "day")) %>%
  left_join(weather, by = "date") %>%
  mutate(datetime_hour = floor_date(datetime, "hour")) %>%
  left_join(rain, by=c("datetime_hour" = "datetime"))
```

### Length

```{r plot_length, echo=F, fig.width=5, fig.height=3, fig.align='center'}

rides.final %>%
  mutate(length.int = cut(log(length), 16)) %>% 
  filter(!is.na(stressful)) %>%
  group_by(length.int) %>%
  summarise(p = mean(stressful), count = n()) %>%
  ggplot(aes(x = length.int, y = p)) + 
  geom_bar(stat="identity") + 
  theme_tufte() + 
  geom_text(aes(label=count, y = p+0.015)) + 
  labs(title = "Distribution of Stressful Rating by Ride Length",
       x = "log length", y = "prob. of stressful rating")

rides.final %>%
  mutate(length.int = cut(log(length), 16),
         stressful.is.na = is.na(stressful)) %>% 
  group_by(length.int) %>%
  summarise(p = mean(stressful.is.na), count = n()) %>%
  ggplot(aes(x = length.int, y = p)) + 
  geom_bar(stat="identity") + 
  theme_tufte() + 
  geom_text(aes(label=count, y = p+0.05)) + 
  labs(title = "Distribution of Nonresponse by Ride Length",
       x = "log length", y = "prob. of nonresponse")

  
```

### Weather

We also want to consider patterns with weather. We have data on daily weather,
including wind speed, temperature highs and lows, and rain data. But we also
have hourly rain data from a local fire station.

```{r plot_weather, echo=F, fig.width=5, fig.height=3, fig.align="center"}

rides.final %>%
  mutate(rainfall.int = cut(log(rainfall.4h+1), 16)) %>% 
  filter(!is.na(stressful)) %>%
  group_by(rainfall.int) %>%
  summarise(p = mean(stressful), count = n()) %>%
  ggplot(aes(x = rainfall.int, y = p)) + 
  geom_bar(stat="identity") + 
  theme_tufte() + 
  geom_text(aes(label=count, y = p+0.015)) + 
  labs(title = "Distribution of Stressful Rating by Recent Rainfall",
       x = "rainfall in past four hours", y = "prob. of stressful rating")

rides.final %>%
  mutate(rainfall.int = cut(log(rainfall.4h + 1), 16),
         stressful.is.na = is.na(stressful)) %>% 
  group_by(rainfall.int) %>%
  summarise(p = mean(stressful.is.na), count = n()) %>%
  ggplot(aes(x = rainfall.int, y = p)) + 
  geom_bar(stat="identity") + 
  theme_tufte() + 
  geom_text(aes(label=count, y = p+0.05)) + 
  labs(title = "Distribution of Nonresponse by Recent Rainfall",
       x = "rainfall in past 4 hours", y = "prob. of nonresponse")

rides.final %>%
  mutate(wind.int = cut(log(mean.wind.speed+1), 16)) %>% 
  filter(!is.na(stressful)) %>%
  group_by(wind.int) %>%
  summarise(p = mean(stressful), count = n()) %>%
  ggplot(aes(x = wind.int, y = p)) + 
  geom_bar(stat="identity") + 
  theme_tufte() + 
  geom_text(aes(label=count, y = p+0.015)) + 
  labs(title = "Distribution of Stressful Rating by Wind Speed",
       x = "daily mean wind speed", y = "prob. of stressful rating")

rides.final %>%
  mutate(wind.int = cut(log(mean.wind.speed+1), 16),
         stressful.is.na = is.na(stressful)) %>% 
  group_by(wind.int) %>%
  summarise(p = mean(stressful.is.na), count = n()) %>%
  ggplot(aes(x = wind.int, y = p)) + 
  geom_bar(stat="identity") + 
  theme_tufte() + 
  geom_text(aes(label=count, y = p+0.05)) + 
  labs(title = "Distribution of Nonresponse by Wind Speed",
       x = "daily mean wind speed", y = "prob. of nonresponse")

```


### Traffic / Daily Trends

We would like to incorporate traffic, but to simplify our model, we may simple
use time of day as a proxy.

```{r plot_traffic, echo=F, fig.width=5, fig.height=3, fig.align='center'}
empirical.p.hourly <- rides.final %>%
  mutate(hour = as.factor(hour(datetime))) %>%
  filter(!is.na(stressful)) %>%
  group_by(hour) %>%
  summarise(p = mean(stressful))

p.na.hourly <- rides.final %>%
  mutate(hour = as.factor(hour(datetime)),
         stressful.is.na = is.na(stressful)) %>%
  group_by(hour) %>%
  summarise(p = mean(stressful.is.na))
  

ggplot(empirical.p.hourly, aes(x = hour, y = p)) +
  geom_bar(stat="identity") + 
  labs(title = "Hourly Probability of Stressful Rating") + 
  theme_tufte()

ggplot(p.na.hourly, aes(x = hour, y = p)) +
  geom_bar(stat="identity") + 
  labs(title = "Hourly Probability of Nonresponse") + 
  theme_tufte()
```



```{r prep_data, echo=FALSE}
# Data Prep
rides.scaled <- rides.final %>%
  mutate(log.length = scale(log(length)),
    length = scale(length)) %>%
  filter(!is.na(stressful),
         !is.na(max.temp),
         !is.na(rain))

# Compute rider average log.length
rider.avg.lengths <- rides.scaled %>%
  group_by(rider) %>%
  summarise(avg.log.length = mean(log.length, na.rm=TRUE))

rides.scaled <- rides.scaled %>%
  left_join(rider.avg.lengths, by="rider")
  
# Create a scaled hour term
rides.scaled <- rides.scaled %>%
  mutate(hour = scale(hour(datetime)))

# Create training and testing sets of rides, stratified by rider
rides.scaled$index <- 1:nrow(rides.scaled)
inTraining <- numeric(0)
riders <- rides.scaled$rider %>% unique()
for (i in riders) {
  their.rides <- rides.scaled %>% 
    filter(rider == i) %>%
    .$index
  rider.ride.indexes <- sample(their.rides, 
                               size = ceiling(0.75 * length(their.rides)))
  inTraining <- c(inTraining, rider.ride.indexes)
}

training <- rides.scaled[inTraining,]
testing <- rides.scaled[-inTraining,]
```
## The Models

### Classical Model

First, we consider how a classical logistic regression model compares to
a model with a random intercept for riders. So we will model:

\[
Y = \text{logit}^{-1} \left(
\alpha + \beta_1 \cdot \text{log.length} + \beta_2 \cdot \text{log.wind speed} +
\beta_3 \cdot \text{log.rainfall.4h}
\right).
\]

### Just Rider Random Effects

Now we want to explore how we can capture variance with and between riders. So
we will use the basic model

\[
Y \sim \text{Bernoulli} (\text{logit}^{-1}(\alpha_{j[i]})),
\quad
\alpha_{j[i]} \sim \text{Normal}(\mu_\alpha, \sigma^2_\alpha).
\]


### Add Time of Day Effects

Now we want to add effects based on time of day. We will try using polynomial
regression to do this first, by adding to our regression the terms,

\[
\beta_1 \cdot \text{hour} + 
\beta_2 \cdot \text{hour}^2 + 
\beta_3 \cdot \text{hour}^3 + 
\beta_4 \cdot \text{hour}^4.
\]

### All Effects

Our last model will take the rider intercepts and day effects and add the terms
we had in our first regression with variables. 


## Table of coefficients

For now, we will compute these models using maximum likelihood. Later, we might
do Bayesian inference with STAN.

```{r fit_models, echo = FALSE}
# Fit each of the models
model1 <- glm(stressful ~ log.length + rainfall.4h + mean.wind.speed + 1,
               data = training, family=binomial(link="logit"))
model2 <- glmer(stressful ~ 1 + (1|rider),
               data = training, family=binomial(link="logit"))
model3 <- glmer(stressful ~ 1 + (1|rider) + 
                  hour + I(hour^2) + I(hour^3) + I(hour^4),
               data = training, family=binomial(link="logit"))
model4 <- glmer(stressful ~ 1 + (1|rider) + 
                  hour + I(hour^2) + I(hour^3) + I(hour^4) + 
                  log.length + rainfall.4h + mean.wind.speed ,
               data = training, family=binomial(link="logit"))
```



```{r create_table, echo=FALSE, results="asis"}
# Now let's make the table of coefficients
stargazer(model1, model2, model3, model4, type="html", title="results",
          align=TRUE)

#stargazer(anova(model2, model3, model4), align=TRUE)
```


### Rider Intercepts

```{r rider_intercepts, echo=F}
rider_intercepts <- data.frame(rider = rownames(coef(model4)$rider[1]),
                           model3 = coef(model3)$rider[1],
                           model4 = coef(model4)$rider[1]) %>%
  tbl_df() %>%
  rename(model3 = `X.Intercept.`, model4=`X.Intercept..1`) %>%
  gather(model, intercept, 2:3)

ggplot(rider_intercepts, aes(x = intercept)) + geom_histogram(bins=30) +
  facet_wrap(~ model, ncol = 1) + 
  theme_tufte() + 
  geom_vline(aes(xintercept = mean(intercept)), color = "red") + 
  labs(title="Distribution of Rider Intercepts")
```

### Hourly Trends

```{r hourly_plot, echo=FALSE, fig.width=5, fig.height=3, fig.align='center'}

empirical.p.hourly.scaled <- rides.scaled %>%
  group_by(hour) %>%
  summarise(p = mean(stressful))

hourly.fun.gen <- function(model) {
  function(hour) {
    coefs <- coef(model)$rider
    intercept <- coefs$`(Intercept)`[1]
    intercept + 
      coefs$hour[1] * hour +
      coefs$`I(hour^2)`[1] * hour^2 +
      coefs$`I(hour^3)`[1] * hour^3 +
      coefs$`I(hour^4)`[1] * hour^4
  }
}
model3.hourly <- hourly.fun.gen(model3)
model4.hourly <- hourly.fun.gen(model4)

hourly.data <- data.frame(hour = seq(min(rides.scaled$hour), max(rides.scaled$hour), length.out=1000))
hourly.data$stress3 <- model3.hourly(hourly.data$hour)
hourly.data$stress4 <- model4.hourly(hourly.data$hour)

hourly.data <- hourly.data %>%
  gather(model, value, stress3, stress4)

ggplot() + 
  geom_line(aes(x = hour, y = invlogit(value), color = model), data = hourly.data) + 
  theme_tufte() + 
  labs(title = "Time of Day Curve",
       x = "Normalized Hour of Day",
       y = "Prob(Stressful)") + 
  geom_bar(aes(x = hour, y = p), data = empirical.p.hourly.scaled, stat="identity")
```


## Model Accuracy and Fit
```{r sep_plot_def, echo=FALSE}
separation.plot <- function(data, col.actual, col.probs) {
  results <- data %>%
    arrange_(col.probs) %>%
    select_(col.actual, col.probs) %>%
    rename_(Y = col.actual, Yhat = col.probs)
  
  expected.true = sum(results$Y)
  
  ggplot(results) +
    geom_rect(aes(xmin = 0, xmax = seq(length.out = length(Yhat)), ymin = 0, ymax = 1),
              fill = "white") +
    geom_linerange(aes(color = Y, ymin = 0, ymax = 1, 
                       x = seq(length.out = length(Yhat)))) + 
    geom_line(aes(y = Yhat, x = seq(length.out = length(Yhat))), lwd = 0.8)  +
    scale_y_continuous("Y-hat\n", breaks = c(0, 0.25, 0.5, 0.75, 1.0)) + 
    scale_x_continuous("", breaks = NULL) +
    theme_linedraw() + 
    scale_colour_grey(start=1, end=0) + 
    geom_point(aes(y = 0, x = length(Yhat) - expected.true), shape=17)
}
```

### Separation Plots

```{r check_new_levels, echo=FALSE}
new.levels.rider <- !setequal(levels(training$rider),
                              levels(testing$rider))

if (new.levels.rider) {
  training <- training %>% filter(rider %in% levels(training$rider))
}
```
Were there new levels? `r new.levels.rider`.
```{r pred_sep_plot, fig.width=5, fig.height=1, echo=FALSE, fig.align='center'}
  
  model.results <- data.frame(stressful = testing$stressful,
                              pred1 = predict(model1, newdata=testing, type="response"),
                              pred2 = predict(model2, newdata=testing, type="response"),
                              pred3 = predict(model3, newdata=testing, type="response"),
                              pred4 = predict(model4, newdata=testing, type="response"))
  
  separation.plot(model.results, "stressful", "pred1")
  separation.plot(model.results, "stressful", "pred2")
  separation.plot(model.results, "stressful", "pred3")
  separation.plot(model.results, "stressful", "pred4")
```

