---
title: "First Model"
author: "Will Jones"
date: "December 8, 2015"
output: pdf_document
header-includes:
- \usepackage{dcolumn}
---

```{r, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(arm))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(rstanarm))
suppressPackageStartupMessages(library(heatmapFit))
suppressPackageStartupMessages(library(stargazer))
suppressPackageStartupMessages(library(gridExtra))
```


## Introduction

As our first steps in modeling ride rating, we will start to model without route
data. Instead we will focus on other question in the modeling as a start for our
model:

- How much variation is there between riders in how they tend to rate rides?
- What relationship does weather, like rain or wind speed, have with ride
rating?
- How does ride rating fluxuate with time of day (which we use as a proxy for
traffic)?

We actually expect a fair amount of the variance in ride rating to be explained
by these variables, based on tests of a smaller sample.

## Some Numbers about the Data

```{r, echo=FALSE}
load('../data/rides.RData')
load('../data/weather_daily.RData')
load('../data/rain_hourly.RData')
```

```{r, echo=FALSE}
total.obs <- nrow(rides)
no.response <- rides %>% filter(is.na(stressful)) %>% nrow()
```

There are `r total.obs` rides in the data set, with `r no.response`
(`r 100 * no.response / total.obs`%) rides with no rating.

## What variables will we include?

```{r, echo=FALSE}
# Merging in weather data
rides.final <- rides %>%
  mutate(date = floor_date(datetime, "day")) %>%
  left_join(weather, by = "date") %>%
  mutate(datetime_hour = floor_date(datetime, "hour")) %>%
  left_join(rain, by=c("datetime_hour" = "datetime"))
```

### Length

```{r, echo=F, fig.width=5, fig.height=3, fig.align='center'}
ggplot(rides.final, aes(x = log(length), fill = stressful)) +
  geom_density(alpha = 0.4) + 
  labs(title = "Ride Rating and Log Length",
       xlab = "Log Length",
       ylab = "Density") + 
  theme_tufte()
```

### Weather

We also want to consider patterns with weather. We have data on daily weather,
including wind speed, temperature highs and lows, and rain data. But we also
have hourly rain data from a local fire station.

```{r, echo=F, fig.width=5, fig.height=3, fig.align="center"}
ggplot(rides.final, aes(x = log(rainfall.4h), fill = stressful)) + 
  geom_density(alpha = 0.4) + 
  labs(title="Ride Ratings and Recent Rainfall",
       xlab="4-Hour Cumulative Rainfall (in)",
       ylab="Density") + 
  theme_tufte()

ggplot(rides.final, aes(x =log(mean.wind.speed), fill = stressful)) + 
  geom_density(alpha = 0.4, adjust = 2.5) + 
  labs(title="Ride Ratings and Daily Windspeed",
       xlab="Mean Wind Speed of Day of Ride)",
       ylab="Density") + 
  theme_tufte()
```


### Traffic / Daily Trends

We would like to incorporate traffic, but to simplify our model, we may simple
use time of day as a proxy.

```{r, echo=F, fig.width=5, fig.height=3, fig.align='center'}
ggplot(rides.final, aes(x = hour(datetime), fill = stressful)) + 
  geom_density(alpha = 0.4) + 
  labs(title = "Ride Rating and Time of Day",
       xlab = "Hour of Day",
       ylab = "Density") +
  theme_tufte()
```



```{r, echo=FALSE}
# Data Prep
rides.scaled <- rides.final %>%
  mutate(log.length = scale(log(length)),
    length = scale(length)) %>%
  filter(!is.na(stressful),
         !is.na(max.temp),
         !is.na(rain))

# Compute rider average log.length
rider.avg.lengths <- rides.scaled %>%
  group_by(rider) %>%
  summarise(avg.log.length = mean(log.length, na.rm=TRUE))

rides.scaled <- rides.scaled %>%
  left_join(rider.avg.lengths, by="rider")
  
# Create a scaled hour term
rides.scaled <- rides.scaled %>%
  mutate(hour = scale(hour(datetime)))


inTraining <- createDataPartition(rides.scaled$id, p = .75, list = FALSE)

training <- rides.scaled[inTraining,]
testing <- rides.scaled[-inTraining,]
```
## The Models

### Intercept Baseline Model

First, we might simply try to model ride rating by modeling ride rating $Y$ as

\[ Y \sim \text{Bernoulli}(p), \]

where $p$ is just the probability that a ride is rated "stressful". Essentially,
what is $p$?

### Classical Model

We also want to consider how a classical logistic regression model compares to
a model with a random intercept for riders. So we will model:

\[
Y = \text{logit}^{-1} \left(
\alpha + \beta_1 \cdot \text{log.length} + \beta_2 \cdot \text{log.windspeed} +
\beta_3 \cdot \text{log.rainfall.4h}
\right).
\]

### Just Rider Random Effects

Now we want to explore how we can capture variance with and between riders. So
we will use the basic model

\[
Y \sim \text{Bernoulli} (\text{logit}^{-1}(\alpha_{j[i]})),
\quad
\alpha_{j[i]} \sim \text{Normal}(\mu_\alpha, \sigma^2_\alpha).
\]


### Add Time of Day Effects

Now we want to add effects based on time of day. We will try using polynomial
regression to do this first, by adding to our regression the terms,

\[
\beta_1 \cdot \text{hour} + 
\beta_2 \cdot \text{hour}^2 + 
\beta_3 \cdot \text{hour}^3 + 
\beta_4 \cdot \text{hour}^4.
\]

### All Effects

Our last model will take the rider intercepts and day effects and add the terms
we had in our first regression with variables. 


## Table of coefficients

```{r, echo = FALSE}
# Fit each of the models
model0 <- glm(stressful ~ 1, data = training, family=binomial(link="logit"))
model1 <- glm(stressful ~ log.length + rainfall.4h + mean.wind.speed + 1,
               data = training, family=binomial(link="logit"))
model2 <- glmer(stressful ~ 1 + (1|rider),
               data = training, family=binomial(link="logit"))
model3 <- glmer(stressful ~ 1 + (1|rider) + 
                  hour + I(hour^2) + I(hour^3) + I(hour^4),
               data = training, family=binomial(link="logit"))
model4 <- glmer(stressful ~ 1 + (1|rider) + 
                  hour + I(hour^2) + I(hour^3) + I(hour^4) + 
                  log.length + rainfall.4h + mean.wind.speed ,
               data = training, family=binomial(link="logit"))
```

```{r, echo=FALSE, results="asis"}
# Now let's make the table of coefficients
stargazer(model0, model1, model2, model3, model4, title="results",
          align=TRUE)
```



## Model Accuracy and Fit
```{r, echo=FALSE}
separation.plot <- function(data, col.actual, col.probs) {
  results <- data %>%
    arrange_(col.probs) %>%
    select_(col.actual, col.probs) %>%
    rename_(Y = col.actual, Yhat = col.probs)
  
  expected.true = sum(results$Y)
  
  ggplot(results) +
    geom_rect(aes(xmin = 0, xmax = seq(length.out = length(Yhat)), ymin = 0, ymax = 1),
              fill = "white") +
    geom_linerange(aes(color = Y, ymin = 0, ymax = 1, 
                       x = seq(length.out = length(Yhat)))) + 
    geom_line(aes(y = Yhat, x = seq(length.out = length(Yhat))), lwd = 0.8)  +
    scale_y_continuous("Y-hat\n", breaks = c(0, 0.25, 0.5, 0.75, 1.0)) + 
    scale_x_continuous("", breaks = NULL) +
    theme_linedraw() + 
    scale_colour_grey(start=1, end=0) + 
    geom_point(aes(y = 0, x = length(Yhat) - expected.true), shape=17)
}
```


```{r, fig.width=5, fig.height=1, echo=FALSE}
model.results <- data.frame(stressful = testing$stressful,
                            pred0 = predict(model0, newdata=testing, type="response"),
                            pred1 = predict(model1, newdata=testing, type="response"),
                            pred2 = predict(model2, newdata=testing, type="response"),
                            pred3 = predict(model3, newdata=testing, type="response"),
                            pred4 = predict(model4, newdata=testing, type="response"))

separation.plot(model.results, "stressful", "pred0")
separation.plot(model.results, "stressful", "pred1")
separation.plot(model.results, "stressful", "pred2")
separation.plot(model.results, "stressful", "pred3")
separation.plot(model.results, "stressful", "pred4")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results="hide", errors=FALSE}
heatmap.fit(model.results$stressful, model.results$pred1, reps=5)
heatmap.fit(model.results$stressful, model.results$pred2, reps=5)
heatmap.fit(model.results$stressful, model.results$pred3, reps=5)
heatmap.fit(model.results$stressful, model.results$pred4, reps=5)
#mean(testing$stressful != testing$prediction1)

```

