---
title: "Exploring Meyers’ Work"
author: "Will Jones"
date: "October 11, 2015"
output: html_document
---

Rennie Meyers, in her Math 241 Final Project did some work with the RIDE data
set, mainly creating maps and exploring the data with different overlays.

Here I want to walk through some of what she did. (Most of the code is adapted
from Meyers' project, which can be found [here](https://github.com/ReedCollegeMATH241/MATH241_PortlandBikeWorks).)

Data sources:

- [Portland Bicycle Network](http://www.civicapps.org/datasets/bicycle-network):
this is from the Portland, OR Civicapps website, and contains all created and 
planned bicycle boulevards, bike lanes, multi-use trails, and signed connections
from February 26, 2010.
- Ride Trips: A geojson file containing several thousand bike trips on the RIDE
app.


```{r, cache=TRUE}
# Load the data
library(RJSONIO)
library(dplyr)
library(ggplot2)
library(rgdal)
library(ggmap)

#load data
PDX <- fromJSON("data/RIDE/trips.json")
```

```{r, cache=TRUE}
##ride data
# Extract information for first ride
orig.coordinates <- PDX[["features"]][[1]]$geometry$coordinates %>% unlist() %>% matrix(ncol=2, byrow=TRUE)
coordinates <- data.frame(
  lat = orig.coordinates[, 1],
  long = orig.coordinates[, 2],
  activity_type = PDX[["features"]][[1]]$properties["activity_type"],
  rating = PDX[["features"]][[1]]$properties["rating"],
  group = 1
)

# Create master data.frame
master.coordinates <- coordinates

# Append information for 2nd thru last ride
for(i in 2:length(PDX[["features"]])){
  orig.coordinates <- PDX[["features"]][[i]]$geometry$coordinates %>% unlist() %>% matrix(ncol=2, byrow=TRUE)
  coordinates <- data.frame(
    lat = orig.coordinates[, 1],
    long = orig.coordinates[, 2],
    activity_type = PDX[["features"]][[i]]$properties["activity_type"],
    rating = PDX[["features"]][[i]]$properties["rating"],
    group = i
  )
  master.coordinates <- bind_rows(master.coordinates, coordinates)
}
```


```{r}
ggplot(master.coordinates, aes(x=long, y=lat, group=group)) + 
  geom_path(alpha=0.3, aes(col=rating), lineend = "butt") +
  scale_color_gradient2(low = "yellow", mid = "green", high = "red") +
  scale_x_continuous(limits=c(-122.7, -122.6)) +
  scale_y_continuous(limits=c(45.475, 45.55)) +
  coord_map() +
  xlab("Longitude") + ylab("Latitude") + 
  ggtitle("RIDE Data for PDX by Rating")  +
  labs(color = "Ride Rating")
```

