---
output: pdf_document
---

```{r include_reedtemplates_1, include = FALSE}
# This chunk ensures that the reedtemplates package is installed and loaded
# This reedtemplates package includes the template files for the thesis and also
# two functions used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")

if(!require(reedtemplates)){
  library(devtools)
  devtools::install_github("ismayc/reedtemplates")
  }
library(reedtemplates)
library(arm)
library(ggplot2)
library(ggthemes)
library(dplyr)
library(gridExtra)
library(ggmap)
```

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.  These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.  Look for the reference to this label at the beginning of Chapter 2.
-->

# Data Sources {#data-sources}

We combine several data sources to do our analysis. Information about
individual rides, including the GPS trace, the rider, and start timestamp were
provided by Ride Report. Weather data were collected from Weather Underground's
archive of the KPDX weather station and a Portland Fire Bureau station.

Our goal in this chapter is to discuss this data and what considerations we
should have in mind before exploring it in depth. This includes how and by whom
the data was collected, who and what is this data actually representative of,
and what samples were taken of the data.

Some of these considerations, such as the limited demographics represented in 
the Ride Report data, pose serious limitations to how our inferences can be
generalized. Others, such the large number of missing responses in the Ride
Report data, motivate the analysis we are doing in this thesis.
Finally, there are other considerations which we will acknowledge here,
but addressing them is out of the scope of this thesis. This data set contains
an abunance of potential research questions, only a fraction of which could be
reasonably addressed in one thesis.

## Ride Report
<!-- who is making the app and why -->
Ride Report's data is the focus of this paper. Knock Software created
the app to collect large amounts of information about urban cyclists' routes
and experiences on those routes. The hope is that this information will be
valuable to city planners.^[Knock's other project is making a cheaper bicycle 
counter for cities to monitor traffic flow, again intended to be sold to cities 
wishing to improve bike infrastructure.]

Ride Report's approach to crowdsourcing this data is particularcly important
to understand. The app automates every piece of the data collection process 
except for the rating given by the rider. Thus, the app casts aside nuanced 
and (somewhat) reliable human input in favor of increasing sample size:
one could imagine a similar app where users have more control over how the
route is recorded, have the ability to rate on a more fine-grained scale, and
are given more direction in what they are rating for. This tradeoff causes two
problems with the reliability with the data.

<!-- how does it work? -->
Before we get into the potential issues in the data collection, though, let's examine
the data collection process itself. When installed on a persons phone, the 
Ride Report app attempts to automatically detect when
the user starts riding their bicycle, based acceleromoter data, when a user
leaves a familiar Wi-Fi network, and some other pieces of information.
When the app detects the start of a ride, it starts recording a GPS trace.
At the end of the users ride, the app detects them getting of their bike (in
a similar process to how it detected the start of a ride) and prompts them to 
give a rating of the ride. The ride data is saved then, even if the user does
not provide a rating.

<!-- potential problems in process -->
<!-- how common are these errors??? -->
This automatic detection of when a ride starts and stops leads to two related and
common errors in the dataset: first, one ride is often split into two or more
rides at points, such as at a stoplight or a train crossing, where a cyclist
stops for an extended period of time; second, car rides are sometimes
misclassified as bicycle rides and vice-versa (car rides are not rated.) The
app allows riders to correct the misclassification, but provides
no way to join split rides back together.

```{r, echo=F, message=F, results='asis'}
label(path = "figure/ride_report_rating.jpg", 
caption = "The Ride Report app's interface has changed significantly between
versions, including the rating text displayed after a ride. This is the current
version as of Februrary 2015.", 
      label = "ride-rating", type = "figure", scale=.2,
options = "tbh")
```

The app only recently became publicly availible and has undergone significant
changes in the course of it's life. In particular, while the ratings have always
been binary, the labels have changed at various points in time. For a while the
rating labels were "Stressful" and "Chill", while now they are labelled 
"Recommend" and "Avoid" (see `r ref("ride-rating", type="figure")`.) Other 
fundamentals of the data collection process, have remained constant.

<!-- who do the rides represent? -->
The data colletion method itself has some problems, but there also may be some
biases in the population of riders using the app. The app is only availible on 
iOS, so only iPhone owners could use this application which may imply a bias 
toward riders of higher socioeconomic status. At the time of the start of the 
thesis, the app was in private beta, meaning only people who actively sought out using
the app were able to use it. Now the application is public and on the Apple 
App Store, making it more widely availible. So many of the earlier rides may be
people within the developer's personal network. Unfortunately, it's hard to make
any solid conclusions about the users of the app because Ride Report doesn't 
collect any demographic data about their riders. 

<!-- data privacy and sampling -->

One other issue with the Ride Report data guided our analysis: privacy.
Because the data involves timestamps and GPS locations of people's commutes, the
data is very sensitive: one could easily infer someone's home and workplace
based on their most commons routes. In fact, this data is protected by an end-user
license agreement (EULA) which prevents sharing of data, without the explicit 
permission of those involved. This presented a logistic challenge: how were we
to do inference and data exploration without access to the data?

By agreement with Knock Software, identifying data must be kept private. With
permission from five riders, Knock was able to give us a small subset consisting
of all the rides from those five riders, to be kept confidential. That is the
data set we used for prototyping models and some basic exploratory analysis. 
Knock also agreed to allow us to run models fitting scripts on larger samples
of their dataset, as long as they were performed on their computers, with
no identifying data leaving their system.

While at first this set up seems like an inconvenience, it actually has some 
advantages. One of the pitfalls of having an entire data set, especially a high 
dimensional one, is that in performing exploratory analyses it is often too
easy to find spurious "statistically significant" results. Instead, we must
come up with our models before running them, greatly limiting the choices we
can make in the garden of forking paths.


<!-- 
TO DO:
- Add map of Portland density of rides
- Add screenshots of rating a ride
- Give graph of missing data
-->


## Weather Data

Slippery roads and formidable winds are no fun for anyone balancing on a
two-wheeled vehicle. Weather is, then, one of the most obvious family of 
predictors for ride rating, at least intuitively. We use the time of a ride to
join in data about the weather conditions during the ride, including

- the temperature,
- whether and how much it is raining,
- whether the roads are wet or have puddles,
- wind and gust speed.

We include the first two, temperature and percipitation, to account for rider
comfort. A swealtering, frigid, or stormy day could make an unpleasant 
experience for a bicyclist and thus could lead to more negative ratings.

On the other hand, we include the last two, wet road and gust speed, as factors
that impact safety. During and after storms, puddles often accumulate in bike
lanes before the center of the road, pushing cyclists into lanes shared by cars,
which are often more dangerous. <!-- citation for this??? -->

Gust speeds impact the aerodynamics of a ride, which are particular important
for bicyclists. It's one of the main reasons cyclists care about getting into
lower (and more aerodynamic) rider positions. Thus, high wind or gust speeds
may affect rider rating.

```{r, results='asis', echo=F, warning=F, message=F}
weather_stations <- data.frame(lat = c(45.5887089, 45.5219685),
                               long = c(-122.5968694, -122.670755),
                               label = c("KPDX Weather", "PFB Rain Guage"))
portland_map <- get_map("portland, oregon",
                        maptype="roadmap",
                        color="bw",
                        zoom=12)
weather_station_map <- ggmap(portland_map) + 
  geom_point(aes(x = long, y = lat), data = weather_stations) + 
  geom_label(aes(x = long - 0.03, y = lat - 0.005, label = label), 
             data = weather_stations,
             nudge_y = -0.007) +
  labs(x = "longitude", y = "latitude")

ggsave("figure/weather_station_map.pdf", weather_station_map, width = 4.5, height = 3.5)

label(path = "figure/weather_station_map.pdf", 
caption = "Positions of weather data collection sites. Daily weather information
was collected at the KPDX weather station at Portland International Airport. 
Hourly percipitation data was collected at the Portland Fire Bureau's rain
guage in downtown Portland.", 
      label = "weather-stations", type = "figure", scale=1,
options = "tbh")
```

We are limiting our study to rides in Portland, Oregon. Given this, we can
first assume that it may be reasonable to expect that riders are used to the
same climate, and thus have somewhat similar responses to weather. This also
makes it reasonable to use data from one nearby weather station, rather than
attempting to collect from several stations and creating a spatial model for
weather.

For daily summaries of weather conditions, we used weather history
from the KPDX weather station at Portland International Airport downloaded from
Weather Underground. From this we were able to get daily weather data, including

- Average, minimum, maximum temperature for the day.
- Total percipitation.
- Mean wind speed, as well as gust speed (speed of brief, strong winds.)

We also got hourly rainfall data from a data stream at the Portland Fire Bureau
Rain Gage at 55 SW Ash St., which is just about the geographic center of
Portland. This just gives raw uncorrected rain guage data, but gives us a fine
grain look at how much rain there has been recently. 

For daily weather data, such as temperature highs and average wind speed, we use
information from the KPDX weather station. This weather station is the best 
calibrated in the area. It is further from the geographic center of the rides
we are examining, but because the weather is daily summary statistics, we
don't expect closer weather stations to be much more informative.
`r ref("weather-stations", type="figure")` shows the geographic positions of
these two stations.


## Notation for the Joined Data Set {#data-notation}

We combined the ride records with weather data by joining by start datestamp of
the ride. We will denote our set of ride-level predictors, each of which is a 
column vector as,
- $x^\text{length}$, ride length, scaled to have mean 0 and standard deviation of 1
- $x^\text{rain}$, rainfall during hour of ride, in tenths of inches
- $x^\text{rain4h}$, rainfall during past four hours before ride, in tenths of inches
- $x^\text{wind}$, mean wind speed for the day, in miles per hour
- $x^\text{gust}$, max gust speed for the day, in miles per hour^[Gust speeds are
the max speeds of winds that are fast, highly variable, and short-term. For METAR
weather stations, which the KPDX station is, gust speeds report the maximum wind
speed when there were rapid fluxations in wind speed with at least 10 knots in 
the difference between the lows and highs.]
- $x^\text{temp}$, average temperature, in degrees Fahrenheit for the day.

We will often represent this set of predictors as the ride-level predictor 
matrix $X = (x^\text{length} \; x^\text{rain} \; x^\text{rain4h} \; x^\text{wind} \;
x^\text{gust}\; x^\text{temp}).$ We also have the predictor $t \in [0, 24]^n$, 
representing the time of day of the ride, measured by hours since midnight. 
(Because of it's special role in building our models, we use the simple notation
of a single letter for it.)

Let $Y_i = 1$ if the $i$th ride received a negative rating, and $Y_i = 0$ if it 
received a positive rating. Choice of coding which events are 0's and which are
1's is arbitrary when making logistic regression models, though we made this 
choice because for the sake of our analysis, negatively rated rides are more
interesting events. For urban planning applications, they define the areas that
need attention.

